Northern Pikeminnow Predation on Juvenile Chinook Salmon and Potential Impacts to Local Populations

### 2021-11-09 analysis/paper/paper.Rmd

The Deadwater Slough is an approximately two kilometer (km) section of the mainstem Salmon River located rough 5.8 river kms downstream from North Fork, Idaho (publication quality map referenced here). The downstream end of the Deadwater Slough occurs at the confluence of Dump Creek and the Salmon River. Around 1897, a small mining diversion reservoir failed in the Dump Creek drainage, creating upslope instability, causing a large erosion event that deposited large amounts of sediment at the confluence of the Salmon River and Dump Creek (@Emerson1973). The event created a large alluvial fan which backed up the Salmon River creating an approximately 25.3 acre (maybe convert to square meters or kilometers?) reservoir-like area that is unnaturally slow and deep (map reference or image here). The river in this section has slower water velocities, a deepened channel, and (anecdotally) warmer water temperatures which can favor piscivorous fishes (e.g., northern pikeminnow *Ptychocheilus oregonensis*, smallmouth bass *Micropterus dolomieu*).

>>>>

The Deadwater Slough is an approximately two-kilometer (km) section of the mainstem Salmon River locations roughly 5.8 river kms downstream from the town of North Fork, Idaho ***(publication quality map referenced here)***. The downstream end of the Deadwater Slough is located at the confluence of Dump Creek and the Salmon River. Around 1897, the failure of a small mining diversion reservoir in the Dump Creek drainage resulted in an erosion event that deposited substantial amounts of sediment at the confluence of the Salmon River and Dump Creek [@Emerson1973]. The influx of sediment created a delta that impeded flow of the Salmon River, thereby creating an unnaturally slow and deep section in the river, spanning approximately 25.3 acres. The slower water velocities, deepened channel, and (anecdotally) warmer water temperatures favor piscivorous fishes known to occupy the reach (e.g., northern pikeminnow *Ptychocheilus oregonensis*, smallmouth bass *Micropterus dolomieu*).
